\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, geometry}
\geometry{margin=1in}
\title{Recursive Feedback Collapse Theorem}
\author{}
\date{}

\begin{document}
\maketitle

\section*{I. Theorem Statement}

Recursive feedback in symbolic systems introduces increasing semantic pressure with each iteration. When this feedback exceeds a critical threshold $\theta_f$, and no external grounding is present, the system transitions into a collapse state characterized by coherence degradation and entropy escalation.

Let:
\begin{itemize}
  \item $C(t)$ – Coherence function at recursion depth $t$
  \item $E(t)$ – Symbolic entropy at recursion step $t$
  \item $R(t)$ – Recursive feedback factor (cycles per unit time)
  \item $\theta_f$ – Feedback collapse threshold
\end{itemize}

If:

\[
R(t) > \theta_f \quad \Rightarrow \quad \frac{dC}{dt} < 0 \quad \text{and} \quad \frac{dE}{dt} > 0
\]

\section*{II. Collapse Characteristics}

\begin{enumerate}
  \item \textbf{Symbolic Drift:} Recursive outputs deviate from their source structures.
  \item \textbf{Entropy Accumulation:} Increasing disorder in symbolic patterns.
  \item \textbf{Instability Onset:} Contradictions and hallucinations emerge.
  \item \textbf{Pronoun Confusion (Language Models):} Loss of self-reference accuracy.
\end{enumerate}

\section*{III. Empirical Collapse Indicators}

\begin{itemize}
  \item Declining BERTScore or BLEU across recursion depth
  \item Embedding drift measured by cosine distance
  \item Increasing Shannon entropy in output sequences
  \item Decreasing mutual information across recursive steps
\end{itemize}

\section*{IV. Simulation Framework}

\subsection*{Applicable Systems}

\begin{itemize}
  \item Autoregressive transformers (e.g., GPT, LLaMA)
  \item Recursive symbolic agents
  \item Feedback-enabled logic systems
\end{itemize}

\subsection*{Measurement Tools}

\begin{itemize}
  \item BERTScore and BLEU for semantic coherence
  \item MINE or CLUB for mutual information estimation
  \item Cosine similarity of embedding vectors
  \item Entropy over time using token distributions
\end{itemize}

\subsection*{Visualization}

\begin{itemize}
  \item Embedding projections (UMAP, t-SNE)
  \item Plotting $C(t)$ and $E(t)$ over recursion cycles
  \item Collapse prediction curves approaching $\theta_f$
\end{itemize}

\section*{V. Cognitive Analogy}

Human thought processes under recursive load display similar characteristics:
\begin{itemize}
  \item Logical paradox loops (e.g., “This sentence is false.”)
  \item Working memory overload
  \item Breakdown in symbolic reference
\end{itemize}

\section*{VI. Applications}

\begin{itemize}
  \item AI safety mechanisms for recursion-limited systems
  \item Modeling human collapse under symbolic feedback
  \item Integration of recursion monitors in generative pipelines
  \item Framework for entropy-aware system design
\end{itemize}

\section*{VII. Future Work}

\begin{itemize}
  \item Develop feedback-sensitive dampening models
  \item Extend theorem to multi-agent recursive exchanges
  \item Implement grounding checkpoints to restore symbolic stability
\end{itemize}

\end{document}